### 2020-04-06 paper/main.tex
init commit

\documentclass[9pt]{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020

% ready for submission
% \usepackage{neurips_2020}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint]{neurips_2020}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2020}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2020}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{bm}
\usepackage{amsmath}
\usepackage{graphicx}

\graphicspath{{./../../plots/}}
\newcommand{\real}{\text{real}}

\title{Bayesian Neural Arithmetic Units -- Leveraging The Complex Plane}
\author{
  Niklas Heim,
  V\'aclav \v Sm\'idl,
  Tom\'a\v s Pevn\'y \\
  Artificial Intelligence Center\\
  Czech Technical University\\
  Prague, CZ 120 00\\
  \texttt{\{niklas.heim, vasek.smidl, tomas.pevny\}@aic.fel.cvut.cz}\\
}

\begin{document}

\maketitle

\begin{abstract}
  Common Neural Networks poorly approximate simple arithmetic operations.  The
  new Neural Arithmetic Units aim to overcome this difficulty, but are limited
  either limited to operate on positive numbers (NALU by
  \citet{trask_neural_2018}), or can only represent simple addition and
  multiplication (NAU & NMU by \citet{madsen_neural_2020}).
  We present the first Neural Arithmetic Unit that operates on the full domain
  of real numbers and is capable of learning arbitrary power functions.
  For a theoretically grounded sparsification of the units we employ Bayesian
  compression on the model weights via Automatic Relevance Determination.
\end{abstract}

\section{Introduction}%
\label{sec:introduction}

It is known that Neural Networks can approximate functions arbitrarily well but
do not necessarily generalize beyond the range inputs that were seen during
training. Neural Arithmetic Units can help to overcome this problem
\cite{trask_neural_2018,madsen_neural_2020}

\section{Introduction}%
\label{sec:introduction}

We fuse the ideas of the \emph{Neural Multiplication Unit} (NMU) by
\citet{madsen_neural_2020} and the \emph{Neural Arithmetic Logic Unit} (NALU)
by \citet{trask_neural_2018}.

The NALU in its original form is capable of addition, subtraction, and
multiplication of \textbf{positive}, high-dimensional vectors. However, for
small inputs convergence is complicated by the use of the logarithm in the
NALU. The logarithm further prevents the NALU of processing negative inputs
correctly.

The NMU fixes both the previous issues, but can only multiply two inputs, and
not represent arbitrary power functions.  Addition is taken care of by the
\emph{Neural Addition Unit} (NAU), which is essentially a dense layer without
bias. A chain of NAU and NMU can thus learn the task defined in
Eq.~\ref{eq:nmu_task}.

\begin{equation}
  \label{eq:nmu_task}
  y = (x_1 + x_2) \cdot (x_1 + x_2 + x_3 + x_4)
\end{equation}

\section{Complex Neural Multiplication Unit}%
\label{sec:complex_neural_multiplication_unit}

Our \emph{complex NMU} (NMUX) can learn arbitrary power functions while still
being able to correctly deal with negative inputs.
The NMUX layer is defined by
\begin{align}
  k_i &= \begin{cases}
     0  & x_i \leq 0 \\
    \pi & x_i > 0
  \end{cases} \\
  \bm r &= |\bm x| \\
  \bm m &= \exp(\bm M \log(\bm r)) \odot \cos(\bm M \bm k)
\end{align}
where $\bm k$ is a vector that is zero where $\bm x$ is positive and $\pi$
where it is negative, $\bm r$ is the elementwise absolute value, and $\bm M$
the multiplication matrix that we want to learn.

The NMUX is inspired by the multiplication part of the NALU, extended to
negative inputs by leveraging taking advantage of the complex logarithm.
The multiplication part of the NALU reads:
\begin{equation}
  \bm m = \exp(\bm M \log(|\bm x| + \epsilon)),
\end{equation}
where $\epsilon$ is a small constant that ensures positive inputs to the real
logarithm.  If use the complex logarithm, this constant becomes superfluous,
but we end up with a complex output in case of negative $x_i$.  In general, for
any complex number $z$\footnote{Note that complex numbers can be represented in
the polar, complex plane where $z=re^{i\varphi}$ with $r>0$ and $\varphi \in
(0,2\pi)$.}
\begin{align}
  \log(z) = \log\left(r\cdot e^{i\varphi}\right)
     = \log(r) + i\varphi.
\end{align}
However, as $z$ is assumed to be real, we can simplify as follows:
\begin{align}
  \log(z) = \log(r) + ik\pi,
\end{align}
where $k \in [0,1]$.

\begin{align}
  \bm m &= \exp(\bm M \log(\bm x)) \\
    &= \exp(\bm M ( \log(\bm r) + i\pi\bm k ))
\end{align}

\begin{align}
  \bm m_{\text{re}} &= \real(\exp(\bm M \log \bm r + i\pi\bm M\bm k)) \\
    &= \real(\exp(\bm M \log \bm r) \odot (\cos(\pi \bm M \bm k) + i \sin(\pi \bm M \bm k))) \\
    &= \exp(\bm M \log \bm r) \odot \cos(\pi \bm M \bm k)
\end{align}

\section{Experimetns}%
\label{sec:experimetns}

\subsection{10 param func}%
\label{sub:10_param_func}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-task-nmu-paper.pdf}
%   \caption{MSE optimization result}%
%   \label{fig:10-param-func-task-nmu-paper}
% \end{figure}
% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-bayes-task-nmu-paper.pdf}
%   \caption{ARD result with MSE starting point}%
%   \label{fig:10-param-func-bayes-task-nmu-paper}
% \end{figure}
% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-bayes-task-nmu-paper-history.pdf}
%   \caption{ARD history}%
%   \label{fig:history}
% \end{figure}
% 
% \subsection{10 param func with sqrt and power}%
% \label{sub:10_param_func_with_sqrt_and_power}
% 
% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-task-sqrt-power.pdf}
%   \caption{10-param-func-task-sqrt-Power}%
%   \label{fig:10-param-func-task-sqrt-power}
% \end{figure}
% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-bayes-task-sqrt-power.pdf}
%   \caption{10-param-func-bayes-task-sqrt-Power}%
%   \label{fig:10-param-func-bayes-task-sqrt-power}
% \end{figure}
% \begin{figure}
%   \centering
%   \includegraphics[width=0.8\linewidth]{10-param-func-bayes-task-sqrt-power-history.pdf}
%   \caption{10-param-func-bayes-task-sqrt-Power-history}%
%   \label{fig:10-param-func-bayes-task-sqrt-power-history}
% \end{figure}

\subsection{Gradient surfaces}%
\label{sub:gradient_surfaces}

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}