### 2020-06-02 papers/main.tex
clean up; get tikz

\documentclass[9pt]{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020

% ready for submission
% \usepackage{neurips_2020}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage{neurips_2020}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2020}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2020}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsthm}         % definitions
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{caption}
\usepackage{mathtools}

\newtheorem*{definition}{Definition}

\usepackage{todonotes}

\usepackage{bm}
\usepackage{amsmath}
\usepackage{graphicx}

% tikz
\usepackage{tikz}
\usetikzlibrary{%
  arrows,%
  backgrounds,
  intersections,
  shapes,
  shapes.misc,% wg. rounded rectangle
  shapes.arrows,%
  chains,%
  matrix,%
  positioning,% wg. " of "
  scopes,%
  decorations.pathmorphing,% /pgf/decoration/random steps | erste Graphik
  shadows,%
  patterns% for hatched rect
}

% plots
\graphicspath{{./../plots/}}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{polar}
\usepgfplotslibrary{smithchart}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{dateplot}

\newcommand{\real}{\text{Re}}
\newcommand{\imag}{\text{Im}}
\newcommand{\wre}{w_{r}}
\newcommand{\wim}{w_{i}}
\newcommand{\Wre}{\bm W_{r}}
\newcommand{\Wim}{\bm W_{i}}
\newcommand{\npu}{\text{NPU}}
\newcommand{\naivenpu}{\text{NaiveNPU}}
\newcommand{\nau}{\text{NAU}}
\newcommand{\nmu}{\text{NMU}}
\newcommand{\nalu}{\text{NALU}}
\newcommand{\inalu}{\text{iNALU}}
\newcommand{\dense}{\text{Dense}}
\newcommand{\chain}{\text{Chain}}

% tables
\usepackage{booktabs}
\renewcommand{\arraystretch}{1.2}

\title{Neural Power Units}
\author{
  Niklas Heim,
  Tom\'a\v s Pevn\'y,
  V\'aclav \v Sm\'idl\\
  Artificial Intelligence Center\\
  Czech Technical University\\
  Prague, CZ 120 00\\
  \texttt{\{niklas.heim, vasek.smidl, tomas.pevny\}@aic.fel.cvut.cz}\\
}

\begin{document}

\maketitle

\begin{abstract}
  Conventional Neural Networks can approximate simple arithmetic operations, but fail
  to generalize beyond the range of numbers that were seen during training.
  \emph{Neural Arithmetic Units} aim to overcome this difficulty, but current
  arithmetic units are either limited to operate on positive numbers or can
  only represent a subset of arithmetic operations. We introduce the \emph{Neural
  Power Unit} (NPU) that operates on the
  \emph{full domain of real numbers} and is capable of \emph{learning
  arbitrary power functions} in a single layer.
  The NPU thus fixes the shortcomings of existing arithmetic units and extends
  their expressivity. This is achieved by using formulas from complex arithmetics
  without requiring a conversion of the network to complex numbers.
  We show where the NPU outperforms its
  competitors in terms of accuracy and sparsity on artificial arithmetic
  datasets. Additionally, we demonstrate how the NPU can be used as a highly
  interpretable model to discover the generating equation of a dynamical system
  purely from the data.
\end{abstract}

\section{Introduction}%
\label{sec:introduction}

Numbers and simple algebra are essential not only to human intelligence but
also to the survival of many other
species~\citep{dehaene_number_2011,gallistel_finding_2018}. A successful,
intelligent agent should therefore be able to perform simple arithmetic.  State
of the art neural networks are capable of learning arithmetic, but they
fail to extrapolate beyond the ranges seen during
training~\citep{suzgun_evaluating_2018,lake_generalization_2018}.  The
inability to generalize to unseen inputs is a fundamental problem that hints at
a lack of \emph{understanding} of the given task. The model merely memorizes
the seen inputs and fails to abstract the true learning task.  The failure of
numerical extrapolation on simple arithmetic tasks has been shown
by~\cite{trask_neural_2018}, who also introduce a new class of \emph{Neural
Arithmetic Units} that show good extrapolation performance on some arithmetic
tasks.

Including Neural Arithmetic Units in common neural networks, promises to
significantly increase their extrapolation capabilities due to their inductive
bias towards numerical computation. This is especially true for tasks in which
the data contains mathematical relationships.  Additionally, they promise to
reduce the amount of parameters that are needed for a given task, and can
therefore drastically improve the explainability of the model.  We demonstrate
how neural arithmetic units as part of a \emph{Neural Ordinary Differential
Equation} (NODE, \cite{chen_neural_2019}) can outperform a dense network with
significantly fewer parameters in Sec.~\ref{sub:fractional_sir_model}.  In
fact, our NPU can be used to directly read out the correct generating ODE from
the fitted model.  This is inline with recent efforts to build
\emph{interpretable} models instead of \emph{explaining} black box models
\citep{rudin_stop_2019}, like conventional neural networks.

The currently available arithmetic units all have different strengths and
weaknesses, but none of them solve simple arithmetic completely.
The \emph{Neural Arithmetic Logic Unit} (NALU) by \cite{trask_neural_2018},
chronologically, was the first arithmetic unit. It can solve
addition ($+$, including subtraction), multiplication ($\times$), and division
($\div$), but is limited to positive inputs. The convergence of the
NALU is quite fragile due to an internal gating mechanism between addition and
multiplication paths as well as the use of a logarithm which is problematic for
small inputs.  Recently, \cite{schlor_inalu_2020} introduced the \emph{improved NALU} (iNALU,
to fix the NALU's shortcomings. However, it significantly increases its complexity and we
observe only a slight improvement in performance.\\
\cite{madsen_neural_2020} solve ($+,\times$) with two new units: the
\emph{Neural Addition Unit} (NAU), and the \emph{Neural Multiplication Unit}
(NMU). Instead of gating between addition and multiplication paths, they are
separate units that can be stacked. They can work with the full range real
numbers, converge much more reliably, but are not capable of representing
division.

\subsection*{Our Contributions}%
\label{sub:our_contribution}

\textbf{Neural Power Unit.}
We introduce a new arithmetic layer (NPU,
Sec.~\ref{sec:neural_power_unit}) which is capable of learning power
functions $(\prod x_{i}^{w_i})$ for arbitrary real input $x_i$ and power $w_i$ thus including multiplication ($x_1\times x_2 = x_1^1 x_2^1$) as well as division $(x_1\div x_2=x_1^{1}x_2^{-1})$ . This is achieved by using formulas from complex arithmetics (Sec.~\ref{sub:naive_npu}).
Stacks of NAUs and NPUs can thus learn the full spectrum of simple arithmetic operations.

\textbf{Convergence improvement.}
We address the known convergence issues of neural arithmetic units by introducing a
\emph{relevance gate} that smooths out the loss surface of the NPU
(Sec.~\ref{sub:npu}). With the relevance gate the NPU reaches extrapolation and
sparsity errors that are on par with the NMU on ($\times$) and outperform NALU
on ($\div,\sqrt{\cdot}$).

\textbf{Interpretability.}
We show how the NPU can be used as a highly interpretable model for discovery of epidemiological models. 
Specifically, we demontrate its ability to provide a model that can be interpreted as fractional SIR model (Sec.~\ref{sub:fractional_sir_model})
that was used to fit the COVID-19 outbreak in various countries
\citep{taghvaei_fractional_2020}.

\section{Related Work}%
\label{sec:related_work}

A number of different approaches to automatically solve arithmetic tasks have
been studied in the recent years. Approaches include Neural GPUs
\citep{kaiser_neural_2016}, Grid LSTMs \citep{kalchbrenner_grid_2016}, Neural
Turing Machines \citep{graves_neural_2014}, and Neural Random Access Machines
\citep{kurach_neural_2016}. They solve tasks like binary addition and
multiplication, or single digit arithmetic. The Neural Status Register
\citep{faber_neural_2020} focusses on control flow.  The Neural Arithmetic
Expression Calculator \citep{chen_neural_2018}, a hierarchical reinforcement
learner is the only method that solves the division problem, but it operates on
character sequences of arithmetic expressions. Related is symbolic integration
with transformers \citep{lample_deep_2019}. Unfortunately, most of the named models
have severe problems with extrapolation \citep{madsen_measuring_2019,saxton_analysing_2019}.
A solution to the extrapolation problem could be Neural Arithmetic Units. They are
designed with an inductive bias towards systematic, arithmetic computation.
However, currently they are limited in their capabilities of expressing the full
range of simple arithmetic operations ($+,\times,\div$).
In the following two sections we briefly describe the currently available
arithmetic layers including their advantages and drawbacks.

\subsection{Neural Arithmetic Logic Units}%
\label{sub:neural_arithmetic_logic_unit}

\citet{trask_neural_2018} have demonstrated the severity of the extrapolation
problem of dense networks for even the simplest arithmetic operations, such as
summing or multiplying two numbers.  In order to increase the power of
abstraction to for arithmetic tasks they propose the \emph{Neural Arithmetic
Logic Unit} (NALU) which is capable of learning ($+,\times,\div$).  However,
the NALU cannot handle negative inputs correctly due to the logarithm in
Eq.~\ref{eq:nalu_mult}:
\begin{definition}[{\bf NALU}]
  The NALU consits of a ($+$) and a ($\times$) path that share their weights $\bm{\hat{W}}$ and $\bm{\hat{M}}$.
\begin{align}
  \label{eq:nalu_add}
  \text{Addition: }       & \bm a = \bm W \bm x
                          & \bm{\hat W}& = \tanh(\bm{W}) \odot \sigma(\bm{M}) \\
  \label{eq:nalu_mult}
  \text{Multiplication: } & \bm m = \exp \bm{\hat W}(\log(|\bm x|+\epsilon)) & &\\
  \text{Output: }         & \bm y = \bm a \odot \bm g + \bm m \odot (1-\bm g) 
                          & \bm g& = \sigma(\bm G\bm x)
\end{align}
\end{definition}
Additionally, the logarithm destabilizes training to an extent that the chance of success
can drop below 20\% for ($+,\times$), it becomes practically impossible to
learn ($\div$), and difficult to learn from small inputs in general
\citep{madsen_measuring_2019}.
\cite{schlor_inalu_2020} provide a detailed description of the shortcomings of
the NALU and they suggest an \emph{improved NALU} (iNALU).  The iNALU addresses
the NALU's problems through a number of mechanisms.  It has independent
addition and multiplication weights for Eq.~\ref{eq:nalu_add} and
Eq.~\ref{eq:nalu_mult}, clips weights and gradients to improve training
stability, regularizes the weights to push them away from zero, and, most
importantly, introduces a mechanism to recover the sign that is lost due to the
absolute value in the logarithm. Additionally, the authors propose to
reinitialize the network if its loss is not improving during training.  We
include the iNALU in one of our experiments and find that it only slightly
improves the NALU's performance (Sec.~\ref{sub:simple_arithmetic_task}) at the
cost of a significantly more complicated unit. Our NPU avoids all these mechanisms
by internally using complex arithmetics.

\subsection{Neural Multiplication Unit \& Neural Addition Unit}%
\label{sub:neural_multiplication_unit}

Instead of trying to fix the NALU's convergence issues
\cite{madsen_neural_2020} propose a new unit for ($\times$) only.
The \emph{Neural Multiplication Unit} (NMU) uses explicit multiplications and learns
to gate between identity and ($\times$) of inputs. The NMU is
defined by Eq.~\ref{eq:nmu} and is typically used in conjunction with the
so-called \emph{Neural Addition Unit} (NAU) in Eq.~\ref{eq:nau}.
\begin{definition}[{\bf NMU \& NAU}]
  NMU and NAU are two units that can be stacked to model ($+,\times$).
\begin{align}
  \label{eq:nmu}
  \text{NMU: } &y_j = \prod_i \hat M_{ij} z_{i} + 1 - \hat M_{ij}  & \hat M_{ij}=\min(\max(M_{ij}, 0), 1)\\
  \label{eq:nau}
  \text{NAU: } &\bm y = \bm{\hat{A}} \bm x & \hat A_{ij}=\min(\max(A_{ij}, -1), 1)
\end{align}
\end{definition}
Both NMU and NAU are regularized with $\mathcal{R} = \sum_{ij} \min(|W_{ij}|,
|1-W_{ij}|)$ and their weights are clipped, which biases them towards learning
an operation or pruning it completely.  The combination of NAU and NMU can thus
learn $(+,\times)$ for both positive and negative inputs. Training NAU and NMU
is stable and succeeds much more frequently than with the NALU, but they cannot
represent ($\div$), which we address with our NPU.

\section{Neural Power Units}%
\label{sec:neural_power_unit}

To fix the deficiencies of current arithmetic units, we propose a new
arithmetic unit (inspired by NALU) that can learn arbitrary power functions $(x^w)$ (including
$\times,\div$) for positive and negative numbers, and still train well.
Combined with the NAU we solve the full range of arithmetic operations.
This is possible through a simple modification of the ($\times$)-path of the
NALU (Eq.~\ref{eq:nalu_mult_2}).  We suggest to use the complex logarithm and
to allow $\bm W$ to become complex.  The complex logarithm is also defined for
negative inputs and a complex $\bm W$ aides convergence (see
Sec.~\ref{sub:fractional_sir_model}). The improvement during training might be
explained by the additional imaginary parameters that make it possible to avoid
regions with an uninformative gradient signal.

\subsection{Naive Neural Power Unit -- NaiveNPU}%
\label{sub:naive_npu}

With the modifications introduced above we can extend the multiplication path
of the NALU from
\begin{align}
  \label{eq:nalu_mult_2}
  \bm m = \exp \bm W(\log_{\text{real}}(|\bm x|+\epsilon))
\end{align}
to use the complex logarithm ($\log\coloneqq\log_{\text{complex}}$) and a
complex weight $\bm W$ to
\begin{align}
  \label{eq:1}
  \bm z = \exp(\bm W\log \bm x) = \exp\left((\bm \Wre + i\Wim) \log\bm x\right).
\end{align}
The complex log in Eq.~\ref{eq:1} lifts the positivity constraint on $\bm x$ resulting
in a layer that can process both positive and negative numbers correctly.  A complex
weight matrix $\bm W$ somewhere in a larger network would result in complex
gradients in other layers.  This would effectively result in doubling the
number of parameters of the whole network. As we are only interested in real
networks outputs, we can avoid this doubling by considering only the real part
of the output $\bm z$:
\begin{align}
  \real(\bm z) &= \real(\exp((\Wre + i\Wim)(\log\bm r + i\pi\bm k))) \\
    \label{eq:npumult}
    &= \exp(\Wre\log\bm r - \pi\Wim\bm k) \odot \cos(\Wim\log\bm r + \pi\Wre\bm k).
\end{align}
Above we have used Euler's formula\footnote{Euler's formula states that for any
real number $x$: $e^{ix} = \cos x + i\sin x$} and the fact that the complex
logarithm for real valued inputs is
\begin{align}
  \log x = \log r + i\theta = \log r + ik\pi,
\end{align}
where $k=0$ if $r\geq0$ and $k=1$ if $r<0$.
A diagram of the NaiveNPU is shown in Fig.~\ref{fig:npu_diagram}.
\begin{definition}[{\bf NaiveNPU}]
  The Naive Neural Power Unit with matrices $\Wre$ and
  $\Wim$ representing real and imaginary part of the complex numbers defined as
  \begin{gather}
    \label{eq:naivenpu_def}
    \bm z = \exp(\Wre \log\bm r - \Wim\bm k) \odot \cos(\Wim\log \bm r + \Wre\bm k), \text{ where }\\
    \nonumber
    \bm r = |\bm x|,
    \quad
    k_i = \begin{cases}
       0  & x_i \leq 0 \\
       \pi & x_i > 0
    \end{cases}
  \end{gather}
\end{definition}

\begin{figure}
  \begin{minipage}{.45\textwidth}
      \resizebox{0.9\textwidth}{!}{\input{../tikz/npu-diagram.tex}}
      \captionof{figure}{NaiveNPU diagram, with input $\bm x$ and output $\bm y$.
      Vectors in green, trainables in orange, functions in blue.}
      \label{fig:npu_diagram}
  \end{minipage}
  \hspace{0.01\textwidth}
  \begin{minipage}{0.54\textwidth}
    \resizebox{\textwidth}{!}{\input{../tikz/gatednpu-diagram.tex}}
    \captionof{figure}{NPU diagram. The NPU has a relevance gate $\bm g$
    (hatched background) in front of the input to the actual unit to prevent
    vanishing gradients.}
    \label{fig:gatednpu_diagram}
  \end{minipage}
\end{figure}

\subsection{The Relevance Gate -- NPU}%
\label{sub:npu}

The NaiveNPU has difficulties to converge on large scale tasks, and to reach
sparse results in cases where the input to a given row is small. We
demonstrate this on a toy example of learning the identity on one of two inputs
(neglecting the second one):
\begin{align}
  \nonumber
  \mathcal{L} &= |\text{m}(x_1,x_2) - x_1|, \quad \text{where}
  & \text{m} = \npu(2,1), \quad
  &x_1 \sim \mathcal U(0,2), \quad x_2 \sim \mathcal U(0,0.05).
\end{align}
The left plot in Fig.~\ref{fig:id_loss} depicts
the the norm of gradient of a batch of two dimensional inputs, one of which is small and
irrelevant. It is clearly visible that even for this simple example, a large
part of the gradient surface of the NaiveNPU is zero.  This vanishing gradient can be explained
as follows. One row of NaiveNPU weights effectively raises each input to a power and
multiplies them: $x_1^{w_1} x_2^{w_2} \dots x_n^{w_n}$. If a single input $x_i$
is constantly close to zero (i.e. irrelevant), the whole row will be zero, no
matter what its weights are and the gradient information on all other weights
is lost.  We therefore introduce a gate on the input of the NPU that can turn
irrelevant inputs into 1s.
A diagram of the NPU is shown in Fig.~\ref{fig:gatednpu_diagram}.
\begin{definition}[{\bf NPU}]
  The NPU extends the NaiveNPU by the relevance gate $\bm g$ on the input $\bm x$.
   \begin{gather}
    \label{eq:gatednpu_def}
    \bm z = \exp(\Wre \log\bm r - \Wim\bm k) 
          \odot \cos(\Wim\log \bm r + \Wre\bm k), \text{ where } \\
    \bm r = \bm{\hat g} \odot |\bm x| + (1-\bm{\hat g}),
    \quad
    k_i = \begin{cases}
       0  & x_i \leq 0 \\
      \hat g_i\pi & x_i > 0
    \end{cases},
    \quad
    \hat g_i = \min(\max(g_i,0),1)
  \end{gather}
\end{definition}
\begin{figure}
  \centering
  \begin{minipage}{0.55\textwidth}
    \includegraphics[width=\textwidth]{npu_gatednpu_id_grad.pdf}
  \end{minipage}
  \begin{minipage}{0.42\textwidth}
  \end{minipage}
  \caption{Absolute value of a gradient of NaiveNPU and NPU for the task
  of learning the identity on $x_1$. Inputs and
  loss are defined on the right, gradient surfaces on the left
  (black areas are beyond the color scale). The correct solution is
  $w_1=1$ and $w_2=0$.  The NaiveNPU has a large zero gradient region for
  $w_2>0.75$, while the NPU's surface is much more informative. The gates
  for central plot are fixed at $g_1=g_2=0.5$ which corresponds to the initial
  gate parameters. During training they will adjust as needed, in this case to
  $g_1=1$ and $g_2=0$. $\bm W_i$ is set to zero in all plots.
  }%
  \label{fig:id_loss}
\end{figure}
The central plot of Fig.~\ref{fig:id_loss} shows the gradient of the NPU
on the identity task with its initial gate setting of $g_1=g_2=0.5$. The
large zero-gradient region of the NaiveNPU is gone. The last plot shows the
same loss for $g_1=1$ and $g_2=0$, which corresponds to the correct gates at
the end of NPU training. The gradient is independent of $w_2$, which
means that it can easily be pruned by a simple regularization such as L1.  In
Sec.~\ref{sub:large_scale_arithmetic_task} we show how important the relevance
gating mechanism is for the convergence and sparsity of large models.  Sparsity
is especially important in order to use the NPU as an interpretable model.

\subsection{Initialization}%
\label{sub:initialization}
We recommend to initialize the NPU with a Glorot Uniform distribution on the
real weights $\Wre$. The imaginary weights $\Wim$ can be initialized to zeros,
so they will only be used where necessary, and the gate $\bm g$ with 0.5, so
the NPU can choose to output 1.
\todo{should this be longer? or not a section at all?}

\section{Experiments}%
\label{sec:experiments}

In Sec.~\ref{sub:fractional_sir_model} we show how the NPU can be used as a
highly interpretable model which can extract the generating equation of an ODE
that contains fractional powers.
In Sec.~\ref{sub:simple_arithmetic_task}\,\&\,\ref{sub:large_scale_arithmetic_task}
we provide experiments on standard arithmetic tasks in which the NPU outperforms
the NALU in all tasks.

\subsection{Equation Discovery of an Epidemiological Model}%
\label{sub:fractional_sir_model}
\todo{other titles: Discovering the Equations of an Epidemic}

Data-driven models such as \emph{SINDy}~\citep{champion_data-driven_2019} or
\emph{Neural Ordinary Differential Equations} (NODE, \cite{chen_neural_2019})
are used more and more in scientific applications.
Recently, \emph{Universal Differential Equations} (UDEs,
\cite{rackauckas_universal_2020}) were introduced which aim to combine
data-driven models with physically informed differential equations to maximize
explainability of the resulting models.
NODEs and UDEs can be tricky to fit, because they run into the same
difficulties as RNNs like vanishing gradients and bifurcations. With this
experiment we want to demonstrate that the NPU works reliably in such settings,
despite the convergence issues that its predecessor (the NALU) has -
even for non-sequential tasks.
If an ODE contains parameters that cannot easily be modelled by dense layers,
the NPU can be a good choice because of its ability to learn any power
function.  Additionally, if interpretability of the result is important, the
NPU can be used recover the generating ODE because of its interpretable design,
instead of trying to explain a black box model in retrospect.

An example of an ODE that contains powers is the fractional SIR model (fSIR,
\cite{taghvaei_fractional_2020}).  The classical SIR model is built from three
quantities of interest: $S$ (susceptible), $I$ (infectious), and $R$
(recovered/removed).  Arguably the most important part of the model is the
transmission rate $r$, which is typically taken to be proportional to the
product of $S$ and $I$.  \citet{taghvaei_fractional_2020} argue that,
especially in the initial phase of an epidemic, the boundary areas of infected
and susceptible cells scale with a fractional power, which leads to
Eq.~\ref{eq:fracsir_r}:
\begin{align}
  \label{eq:fracsir}
  \frac{dS}{dt} &= -r(t) + \eta R(t), &
  \frac{dI}{dt} &=  r(t) - \alpha I(t), &
  \frac{dR}{dt} &= \alpha I(t) - \eta R(t),\\
  \label{eq:fracsir_r}
  & & r(t) &= \beta I(t)^\gamma S(t)^\kappa,
\end{align}
We solve the fSIR model numerically with the parameters $\alpha=0.05$,
$\beta=0.06$, $\eta=0.01$, $\gamma=\kappa=0.5$, and 40 datapoints that are
equally spaced in the time interval $T=(0,200)$. The initial conditions are set
to $S_0=100$, $I_0=0.01$, and $R_0=0$.

We fit the fSIR model with three different model types: a dense network, the
NPU, and an NPU with imaginary weights fixed to zero (further referred to as
\emph{real NPU}). An exemplary model is: $\npu=\chain(\npu(3,h),\nau(h,3))$
with variable hidden size $h$.  The detailed models are defined in
Tab.~\ref{tab:fsir_models}. The training objective is the loss $\mathcal{L}$
with L1 regularization.
\begin{align}
  \label{eq:sir_loss}
  \mathcal{L} = \text{MSE}(\bm t,\text{model}(\bm u_0)) + \beta||\bm\theta||_1.
\end{align}
We train each model for 3000 steps with the ADAM optimizer and a learning rate
of 0.005, and after that with LBFGS until convergence (or for maximum 1000
steps).  For each model type we run a small grid search to build a Pareto front
with $h\in \{6,9,12,15,20\}$ and $\beta\in\{0,0.01,0.1,1\}$, where each
hyper-parameter pair is run five times. The resulting Pareto front is shown on
the left of Fig.~\ref{fig:pareto_sir}. The NPU clearly reaches much sparser and
better solutions than the dense network.
The real NPU has problems to converge in the majority of cases.
However, there are a few models in the bottom left that
reach a very low MSE and have very few parameters. The best these models is
shown in Fig.~\ref{fig:sir_gatednpu_modelps}.  It looks strikingly similar to
the fSIR model in matrix from:
\renewcommand*{\arraystretch}{1.3}
\begin{equation}
  \begin{bmatrix}
    \dot S \\ \dot I \\ \dot R
  \end{bmatrix}
  =
  \begin{bmatrix}
    -\beta & 0 & \eta \\
    \beta & -\alpha & 0 \\
    0 & \alpha & \eta
  \end{bmatrix}
  \begin{bmatrix}
    I^\gamma S^\kappa \\ I \\ R
  \end{bmatrix}.
\end{equation}
Reading Fig.~\ref{fig:sir_gatednpu_modelps} from right to left, we can extract
the ODE that the real NPU represents.  It identified the transmission rate
correctly as a product of two fractional powers $r = I^\gamma S^\kappa$ with
$\kappa=0.57$ and $\gamma=0.62$, which is close to the true values
$\gamma=\kappa=0.5$.  In the second layer the NAU combines the correct hidden
outputs from the NPU such that $\dot S$ is composed of the negative
transmission rate $r$ and positive $R$.  $\dot I$ and $\dot R$ are also
composed from the correct hidden variables, with the parameters
$\alpha,\beta,\eta$ being not far off from the truth.
We conclude that even with this very naive approach the real NPU is able to
recover a fractional SIR model.
\begin{figure}
  \centering
  \resizebox{.7\textwidth}{!}{\input{../plots/pareto_sir.tex}}
  \caption{Pareto fronts of the dense network, the NPU, and the NPU
  with imaginary weights fixed to zero. The NPU reaches solutions with lower
  MSE and fewer parameters than the dense net. The real NPU mostly yields
  worse results than the NPU, just in a few cases it converges to very sparse
  models with good MSE.}%
  \label{fig:pareto_sir}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=.6\linewidth]{../plots/sir_gatednpu_modelps.pdf}
  \captionof{figure}{Visualization of the best real NPU. Reading from right
  to left, it takes the SIR variables as an input, then applies the NPU and
  the NAU.  It correctly identifies $r$ as a fractional product in the NPU,
  and gets the rest of the fSIR parameters almost right in the NAU.}%
  \label{fig:sir_gatednpu_modelps}
\end{figure}
\\
In summary, the NPU clearly works well in sequential tasks, and we have shown
that we can reach highly interpretable results with the real NPU, but in practice,
using the real NPU might be difficult due to its lower success rate. With a more
elaborate analysis it might be possible to reach the same solutions with the
full NPU and e.g. a strong regularization of its imaginary parameters.

\subsection{Simple Arithmetic Task}%
\label{sub:simple_arithmetic_task}

In this experiment we demonstrate the capabilities of five different
layers (NPU, NMU, NALU, iNALU, Dense) on a small problem with two inputs
and four outputs. The objective is to learn the function $f: \mathbb{R}^2
\rightarrow \mathbb{R}^4$ with a standard MSE loss:
\begin{align}
  \label{eq:simple_f}
  f(x,y) &= (x+y,\, xy,\, x/y,\, \sqrt{x} \text{  })^T \eqqcolon \bm t, \\
  \mathcal L &= \frac{1}{4} \sum_{i=1}^{n=4}\left(\text{model}(x,y)_i - f(x,y)_i\right)
             = \text{MSE}(\bm{\hat t}, \bm t).
\end{align}
Each model has two layers with a hidden dimension $h=6$.  E.g. the NPU model is
defined by $\npu=\chain(\npu(2,6),\nau(6,4))$.  The remaining models that are
used in the tables and plots are given in Tab.~\ref{tab:models_simple_task}.
To obtain valid results in case of division we train on positive, non-zero
inputs, but test on negative, non-zero numbers (except for test inputs to the
square-root):
\begin{align}
  \label{eq:simple_train}
  (x_{\text{train}}, y_{\text{train}}) &\sim \mathcal{U}(0.1,2) &
  (x_{\text{test}}, y_{\text{test}}) &\sim \mathcal{R}(\text{-4.1:0.2:4}) &
  (x_{\text{test,\,sqrt}}, y_{\text{test,\,sqrt}}) &\sim \mathcal{R}(\text{0.1:0.1:4})
\end{align}
where $\mathcal{R}$ denotes a \emph{range} with start, step, and end.
We train each model for 20\,000 steps with the ADAM optimizer, a learning rate of
0.001, and a batch size of 100. The input samples are generated on the fly
during training. Fig.~\ref{fig:simple_err} shows the error surface of the best
of 20 models on each task. Tab.~\ref{tab:simple_err} lists the corresponding
averaged testing errors of all 20 models.

The NPU successfully learns ($+,\times,\div,\sqrt{\cdot}$) and clearly
outperforms NALU and iNALU on all tasks.  The NPU is on par with the NMU for
($+$), but the NMU is better at ($\times$) due to its inductive bias. The NMU
cannot learn ($\div,\sqrt{\cdot}$). The NPU excels at extrapolation on all tasks.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{simple_err.pdf}
  \caption{Comparison of different models learning Eq.~\ref{eq:simple_f}. Each
  column represents the best model of 20 runs.  The dense network learns all
  tasks, but fails to extrapolate. NALU and iNALU are better at extrapolating
  to positive numbers, but fail if one of the inputs is negative.  The NMU
  successfully learns ($+,\times$), but not ($\div,\sqrt{\cdot}\,$). Only the NPU
  learns all tasks and can extrapolate both to positive and negative numbers.
  }%
  \label{fig:simple_err}
\end{figure}

\subsection{Large Scale Arithmetic Task}%
\label{sub:large_scale_arithmetic_task}

One of the most important properties of a layer in a neural network is its
ability to scale. With the large scale arithmetic task we show that the NPU
works reliably on many-input tasks that are heavily over-parametrized.  In this
section we compare NALU, NMU, NPU, and the NaiveNPU on a task that is identical
to the `arithmetic task' that \cite{madsen_neural_2020} and
\cite{trask_neural_2018} analyse as well.
The goal is to sum two subsets of a 100 dimensional vector and
apply an operation (like $\times$) to the two summed subsets.  The dataset
generation is defined in the set of Eq.~\ref{eq:large_scale_def}, with the
parameters from Tab.~\ref{tab:dataset_params_large_scale}.
\begin{align}
  \label{eq:large_scale_def}
  a = \sum_{i=s_{1,\text{start}}}^{s_{1,\text{end}}} x_i, &&
  b = \sum_{i=s_{2,\text{start}}}^{s_{2,\text{end}}} x_i, &&
  t_{\text{add}} = a + b, && t_{\text{mul}} = a \times b, &&
  t_{\text{div}} = 1 / a, && t_{\text{sqrt}} = \sqrt{a},
\end{align}
where starting and
ending values $s_{i,\text{start}},s_{i,\text{end}}$ of the summations are chosen
such that $a$ and $b$ come from subsets of the input vector $\bm x$ with a
given overlap.  Specifics of the used models are defined in
Tab.~\ref{tab:models_large_task}.  \cite{madsen_neural_2020} perform an
extensive analysis of this task with different subset and overlap ratios,
varying model and input sizes, and much more, establishing that the combination
of NAU/NMU outperforms the NALU. We focus on the
comparison of NPU, NMU, and NALU on the default parameters of
\cite{madsen_neural_2020} which sets the subset ratio to 0.5 and the overlap
ratio to 0.25 (details in Tab.~\ref{tab:dataset_params_large_scale}).
We also compare to the NaiveNPU (without the relevance gate) to show
how important the gating mechanism is for both sparsity and overall
performance.

Fig.~\ref{fig:pareto} plots training and testing errors over the number of
non-zero parameters for all models and tasks.
The addition column shows that all models except NaiveNPU learn ($+$) on the
training set, with the NMU converging to the sparsest models. The NMU also
performs best on the testing set, with the NPU as a close second.\\
On ($\times$), the best NMU models also outperform the NPU, but some of
them do not converge at all, which is why the NPU has a better testing MSE on
average (see also Tab.~\ref{tab:arithmetic100_val}). The testing MSE of the
NALU is so large that is excluded from the plot.\\
On ($\div,\sqrt{\cdot}$) the NPU clearly outperforms all other layers
in MSE and sparsity. Generally, the difference between the NaiveNPU and the NPU
is very significant and demonstrates how important the relevance gate is both
for convergence and sparsity. The NPU effectively converts irrelevant inputs
to 1s, while the NaiveNPU is stuck on the zero gradient plateau.

\begin{figure}
  \centering
  \resizebox{\textwidth}{!}{\input{../plots/pareto.tex}}
  \caption{Validation MSE over number of non-zero parameters ($w_i > 0.001$) of the large scale
  arithmetic task.  Again, the NMU outperforms the NPU on its
  native tasks, addition and multiplication. The NPU is the best at division
  and square-root.  The NaiveNPU without the relevance gate is far off, because
  it does not have the necessary gradient signal to converge, as discussed in
  Sec.~\ref{sub:npu}}%
  \label{fig:pareto}
\end{figure}

\begin{table}
  \centering
  \caption{Testing errors of the large scale arithmetic task.}
  \label{tab:arithmetic100_val}
  \input{../plots/arithmetic100_val.tex}
\end{table}

\section{Conclusion}%
\label{sec:conclusion}

We introduced the \emph{Neural Arithmetic Unit} that addresses the deficiencies
of current arithmetic units: it can learn arbitrary power functions for
positive, negative, and small numbers. We showed that the NPU outperforms its
main competitor (NALU) and reaches performance that is on par with the
multiplication specialist NMU.  For tasks that are known to contain only
addition and multiplication a stack of NAU and NMU can be a better choice than
the NPU, because they might be easier to train. For more complicated problems
that involve division or power functions, the NPU is the better choice.

Additionally we have demonstrated how the NPU can be used as an interpretable
model that can recover governing equations of dynamical systems purely from
data.

\section{Statement of Broader Impact}%
\label{sec:statement_of_broader_impact}

Current neural network architectures can be problematic because they are black
box models that are difficult to interpret. This becomes highly problematic if
ML models are involved in high stakes decisions in e.g. criminal justice,
healthcare, or control systems that affect lives.  With the NPU we hope to
contribute to the broad topic of interpretable machine learning, with a focus
on scientific applications.

Additionally, learning to abstract (mathematical) ideas and extrapolate is a
fundamental goal that might contribute to more reliable machine learning
systems.

\section{Acknowledgements}%
\label{sec:acknowledgements}

We thank the authors of the Julia packages
Flux.jl~\citep{innes_fashionable_2018} and
DifferentialEquations.jl~\citep{rackauckas_differentialequationsjl_2017}).

\bibliographystyle{plainnat}
\bibliography{refs}

\appendix
\section*{Appendix}%
\label{sec:appendix}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\begin{table}[h]
  \centering
  \caption{Validation error of the different models (i.e. mean of each heatmap
  in Fig.~\ref{fig:simple_err}). Each value is obtained by averaging the error
  of 20 models.
  }
  \label{tab:simple_err}
  \input{../plots/simple_err.tex}
\end{table}

\begin{table}[h]
  \centering
  \caption{Model definitions for the simple arithmetic task}
  \label{tab:models_simple_task}
  \begin{tabular}{llll}
    \toprule
    Model & Layer 1 & Layer 2 & Layer 3 \\
    \midrule
    NPU & NAU(2,\,6) & NPU(6,\,2) & -- \\
    NMU & NAU(2,\,6) & NMU(6,\,2) & -- \\
    NALU & NALU(2,\,6) & NALU(6,\,2) & -- \\
    iNALU & iNALU(2,\,6) & iNALU(6,\,2) & -- \\
    Dense & Dense(2,\,10,\,\sigma) & Dense(10,\,10,\,\sigma) & Dense(10,\,2) \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{Model definitions for the large scale arithmetic task}
  \label{tab:models_large_task}
  \begin{tabular}{llll}
    \toprule
    Model & Layer 1 & Layer 2\\
    \midrule
    NPU & NAU(100,\,100) & NPU(100,\,1) \\
    NPU & NAU(100,\,100) & NPU(100,\,1) \\
    NMU & NAU(100,\,100) & NMU(100,\,1) \\
    NALU & NALU(100,\,100) & NALU(100,\,1) \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \centering
  \caption{Dataset parameters for the large scale arithmetic task.}
  \label{tab:dataset_params_large_scale}
  \begin{tabular}{lccccc}
    \toprule
    Task & Input size & Subset ratio & Overlap ratio & Training range  & Validation range\\
    \midrule
    Add  & 100        & 0.5          & 0.25          & Sobol(-1,1)     & Sobol(-4,4) \\
    Mult & 100        & 0.5          & 0.25          & Sobol(-1,1)     & Sobol(-4,4)\\
    Div  & 100        & 0.5          & --            & Sobol(0,0.5)    & Sobol(-0.5,0.5) \\
    Sqrt & 100        & 0.5          & --            & Sobol(0,2)      & Sobol(0,4) \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \centering
  \caption{fSIR model definitions}
  \label{tab:fsir_models}
  \begin{tabular}{llll}
    \toprule
    Model & Layer 1 & Layer 2 & Layer 3 \\
    \midrule
    NPU & $\npu(3,\,h)$ & $\nau(h,\,3)$ & -- \\
    NPU & NPU$_{\text{real}}(3,\,h)$ & $\nau(h,\,3)$ & -- \\
    Dense & $\dense(2,\,h,\sigma)$ & $\dense(h,\,h,\,\sigma)$ & $\dense(h,\,3)$ \\
    \bottomrule
  \end{tabular}
\end{table}

\end{document}