### 2020-06-02 plots/pareto.tex
plots

\begin{tikzpicture}[/tikz/background rectangle/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, draw opacity={1.0}}, show background rectangle]
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(1.02, 1)}, anchor={north west}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={1.0mm}, yshift={-1.0mm}, width={55.15mm}, height={48.8mm}, scaled x ticks={false}, xlabel={Nr. Parameters}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={76.74464101797544}, xmax={14403.350974579345}, xtick={{100.0,316.22776601683796,1000.0,3162.2776601683795,10000.0}}, xticklabels={{$10^{2.0}$,$10^{2.5}$,$10^{3.0}$,$10^{3.5}$,$10^{4.0}$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={Testing MSE}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={0.0005672734950819899}, ymax={291455.14136474277}, ytick={{0.01,1.0,100.0,10000.0}}, yticklabels={{$10^{-2}$,$10^{0}$,$10^{2}$,$10^{4}$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.0;green,0.6056;blue,0.9787}, name path={8432bdbe-b384-493b-a894-2ca9aed19dd4}, draw opacity={0.7}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.0;green,0.6056;blue,0.9787}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (1212.0,0.0044327993)
            (633.0,0.025323598)
            (2198.0,0.07513304)
            (1414.0,0.07950895)
            (1045.0,0.08757629)
            (423.0,0.09544645)
            (172.0,0.12249249)
            (1131.0,0.12339931)
            (976.0,0.148304)
            (1970.0,0.14878139)
        }
        ;
    \addplot[color={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, name path={19c422f2-2c28-4773-bff4-dbd0b472008a}, draw opacity={0.7}, line width={0}, solid, mark={pentagon*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (10169.0,398.8512385486777)
            (10217.0,416.529448488352)
            (10157.0,429.61028780569137)
            (12420.0,487.83743759498134)
            (10159.0,621.4133164105747)
            (10246.0,858.7089297238343)
            (11411.0,1077.5153824358792)
            (10300.0,1486.1756888766222)
            (10184.0,1651.9431386116044)
            (11284.0,9942.065087422989)
        }
        ;
    \addplot[color={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, name path={2a7cf6d3-80f4-4533-929d-ea31232589ab}, draw opacity={0.7}, line width={0}, solid, mark={diamond*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (4109.0,0.0010007463)
            (89.0,0.0036541214)
            (90.0,0.0039218063)
            (89.0,0.005820834)
            (90.0,0.00598223)
            (189.0,0.006059413)
            (91.0,0.006182335)
            (90.0,0.006198649)
            (3228.0,0.006199132)
            (90.0,0.043522302)
        }
        ;
    \addplot[color={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, name path={b82a4fdf-f8e3-4a56-bc52-e5e1e40eaa88}, draw opacity={0.7}, line width={0}, solid, mark={triangle*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (9878.0,161.49628)
            (9879.0,161.5955)
            (9852.0,161.59552)
            (9837.0,161.59552)
            (9836.0,161.59557)
            (9944.0,161.70662)
            (9854.0,186.75471)
            (9861.0,250.42944)
            (9933.0,388.22473)
            (9870.0,165211.48)
        }
        ;
\end{axis}
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(1.02, 1)}, anchor={north west}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={58.15mm}, yshift={-1.0mm}, width={55.15mm}, height={48.8mm}, scaled x ticks={false}, xlabel={Nr. Parameters}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={89.49402359104843}, xmax={35417.024208050476}, xtick={{100.0,316.22776601683796,1000.0,3162.2776601683795,10000.0,31622.776601683792}}, xticklabels={{$10^{2.0}$,$10^{2.5}$,$10^{3.0}$,$10^{3.5}$,$10^{4.0}$,$10^{4.5}$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={0.1}, ymax={100000.0}, ytick={{0.1,1.0,10.0,100.0,1000.0,10000.0,100000.0}}, yticklabels={{$10^{-1}$,$10^{0}$,$10^{1}$,$10^{2}$,$10^{3}$,$10^{4}$,$10^{5}$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.0;green,0.6056;blue,0.9787}, name path={ca743fe9-f865-4257-9e33-032174efb577}, draw opacity={0.7}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.0;green,0.6056;blue,0.9787}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (1947.0,3.119196)
            (1912.0,3.1269026)
            (1934.0,3.200298)
            (1314.0,3.550602)
            (2021.0,3.9671524)
            (2036.0,4.5925226)
            (2517.0,4.6573997)
            (1865.0,4.9660153)
            (1928.0,7.385417)
            (2304.0,8.159821)
        }
        ;
    \addplot[color={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, name path={718cbf1c-0e6b-42da-8705-e6a8efb5b5de}, draw opacity={0.7}, line width={0}, solid, mark={pentagon*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (27266.0,6.056448979580512e57)
            (27271.0,1.1270697569661482e67)
            (29240.0,6.670123662775398e74)
            (29769.0,2.3908233870509967e81)
            (29902.0,5.73026454938337e83)
            (29723.0,5.578735868305779e84)
            (29719.0,6.47201021019638e89)
            (29857.0,3.343027269532066e95)
        }
        ;
    \addplot[color={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, name path={3adb2d3b-7fa3-4ded-a575-cc69f8a5d964}, draw opacity={0.7}, line width={0}, solid, mark={diamond*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (1432.0,0.22429785)
            (3584.0,0.35476318)
            (106.0,0.50807154)
            (5648.0,0.5131744)
            (109.0,0.597559)
            (8249.0,2.7305448)
            (8075.0,20011.611)
            (383.0,27771.52)
            (383.0,27781.611)
            (385.0,27991.143)
        }
        ;
    \addplot[color={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, name path={1b311080-2e2a-4c14-b593-bbd939282850}, draw opacity={0.7}, line width={0}, solid, mark={triangle*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (9938.0,687.11816)
            (9950.0,1251.8556)
            (9983.0,2757.291)
            (9962.0,3324.5964)
            (10018.0,4169.526)
            (9966.0,4312.167)
            (10049.0,4416.4814)
            (10071.0,4609.3774)
            (10056.0,4618.8584)
            (9946.0,5408.8984)
        }
        ;
\end{axis}
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(1.02, 1)}, anchor={north west}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={115.3mm}, yshift={-1.0mm}, width={55.15mm}, height={48.8mm}, scaled x ticks={false}, xlabel={Nr. Parameters}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={43.480244741743846}, xmax={24215.503069354898}, xtick={{100.0,316.22776601683796,1000.0,3162.2776601683795,10000.0}}, xticklabels={{$10^{2.0}$,$10^{2.5}$,$10^{3.0}$,$10^{3.5}$,$10^{4.0}$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={3.003128636465643e-10}, ymax={1.0476863806782078e31}, ytick={{1.0,1.0e10,1.0e20,1.0e30}}, yticklabels={{$10^{0}$,$10^{10}$,$10^{20}$,$10^{30}$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.0;green,0.6056;blue,0.9787}, name path={25c6bfe2-b14c-4d58-812c-34611c63a621}, draw opacity={0.7}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.0;green,0.6056;blue,0.9787}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (52.0,4.217039e-9)
            (155.0,4.6643503e-8)
            (204.0,5.3117162e-8)
            (102.0,5.7346426e-8)
            (102.0,9.8668096e-8)
            (154.0,1.7296065e-7)
            (102.0,2.6649218e-7)
            (102.0,3.3172265e-7)
            (102.0,3.909838e-7)
            (102.0,4.380334e-7)
        }
        ;
    \addplot[color={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, name path={35cbff1c-ba07-48c9-bd21-cbf564ba8620}, draw opacity={0.7}, line width={0}, solid, mark={pentagon*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (18994.0,0.006679143442147095)
            (20248.0,300.5367981426397)
            (18253.0,420.4417069693)
            (18484.0,462.16137066417264)
            (19124.0,506.4912807643721)
            (19352.0,550.5197097967591)
            (18115.0,699.409735838578)
            (19284.0,806.9032735843994)
            (18698.0,1548.9257860967384)
            (19035.0,2074.445574153203)
        }
        ;
    \addplot[color={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, name path={364b3eb2-628a-44de-a4b5-f6fca4d5743c}, draw opacity={0.7}, line width={0}, solid, mark={diamond*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (801.0,1.3877087)
            (749.0,1.5402892)
            (757.0,1.5402905)
            (742.0,1.5402931)
            (757.0,1.5402986)
            (692.0,1.7030367)
            (693.0,1.7030562)
            (864.0,1.7975432)
            (791.0,1.9803718)
            (715.0,2.169463)
        }
        ;
    \addplot[color={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, name path={e362d5ba-5d53-4952-9528-3385a563c01a}, draw opacity={0.7}, line width={0}, solid, mark={triangle*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (9845.0,2.0357272e7)
            (9810.0,4.9806867e8)
            (9970.0,6.223722e13)
            (10082.0,7.144208e14)
            (9868.0,7.4999905e15)
            (9765.0,1.0666438e18)
            (9927.0,2.085579e20)
            (9760.0,7.4610096e29)
        }
        ;
\end{axis}
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(1.02, 1)}, anchor={north west}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={172.45mm}, yshift={-1.0mm}, width={55.15mm}, height={48.8mm}, scaled x ticks={false}, xlabel={Nr. Parameters}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={86.16627370302126}, xmax={33416.29939718559}, xtick={{100.0,316.22776601683796,1000.0,3162.2776601683795,10000.0,31622.776601683792}}, xticklabels={{$10^{2.0}$,$10^{2.5}$,$10^{3.0}$,$10^{3.5}$,$10^{4.0}$,$10^{4.5}$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={0.0014075044814322237}, ymax={2.162048835278372e21}, ytick={{1.0,100000.0,1.0e10,1.0e15,1.0e20}}, yticklabels={{$10^{0}$,$10^{5}$,$10^{10}$,$10^{15}$,$10^{20}$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.0;green,0.6056;blue,0.9787}, name path={ebdc7105-4d92-4a3f-95d2-02699858279c}, draw opacity={0.7}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.0;green,0.6056;blue,0.9787}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (103.0,0.006807239)
            (103.0,0.036651038)
            (103.0,0.040400617)
            (103.0,0.052491613)
            (102.0,0.053030577)
            (102.0,0.054934423)
            (103.0,0.05519525)
            (104.0,0.05592795)
            (103.0,0.27859107)
            (103.0,0.28152826)
        }
        ;
    \addplot[color={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, name path={554d6241-181a-4734-8d6d-49428305bc1c}, draw opacity={0.7}, line width={0}, solid, mark={pentagon*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (25002.0,110.9731299000638)
            (24711.0,139.27110419326715)
            (23060.0,259.65656416957603)
            (24543.0,883.9527082423043)
            (23574.0,2017.814681341383)
            (28229.0,12599.269355463382)
            (27241.0,19142.912878289713)
            (20686.0,42905.89328481797)
            (27171.0,91624.4060802308)
            (20707.0,3.556283999184454e7)
        }
        ;
    \addplot[color={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, name path={55056117-138d-4bdc-a592-4d4fe5a3384e}, draw opacity={0.7}, line width={0}, solid, mark={diamond*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (7201.0,9.486513)
            (6707.0,9.52155)
            (188.0,9.844871)
            (6927.0,9.845515)
            (1701.0,10.554757)
            (1101.0,11.371628)
            (10001.0,11.378366)
            (5902.0,11.498858)
            (8701.0,11.621315)
            (10001.0,13.208106)
        }
        ;
    \addplot[color={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, name path={f886d280-b26f-49cb-9603-d33a459d2008}, draw opacity={0.7}, line width={0}, solid, mark={triangle*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.7}, fill={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, fill opacity={0.7}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (7992.0,0.23132631)
            (9564.0,99.999794)
            (7554.0,99.999794)
            (8926.0,100.0038)
            (7772.0,100.0563)
            (4221.0,1.8681482e9)
            (6967.0,3.5018813e10)
            (6613.0,1.2335086e12)
            (6682.0,4.4703785e20)
        }
        ;
\end{axis}
\end{tikzpicture}