### 2020-06-02 plots/pareto_sir.tex

\begin{tikzpicture}[/tikz/background rectangle/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, draw opacity={1.0}}, show background rectangle]
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={Pareto}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(0.98, 0.98)}, anchor={north east}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={1.0mm}, yshift={-1.0mm}, width={112.3mm}, height={74.2mm}, scaled x ticks={false}, xlabel={Nr. Paramters}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={15.305486448632008}, xmax={625.3313171144225}, xtick={{15.848931924611142,25.11886431509581,39.810717055349734,63.09573444801933,100.0,158.48931924611142,251.18864315095823,398.1071705534973}}, xticklabels={{$10^{1.2}$,$10^{1.4}$,$10^{1.6}$,$10^{1.8}$,$10^{2.0}$,$10^{2.2}$,$10^{2.4}$,$10^{2.6}$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={MSE}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={0.005}, ymax={10000.0}, ytick={{0.01,0.1,1.0,10.0,100.0,1000.0,10000.0}}, yticklabels={{$10^{-2}$,$10^{-1}$,$10^{0}$,$10^{1}$,$10^{2}$,$10^{3}$,$10^{4}$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, name path={c6416611-1574-49a0-9ea4-91e50a33118b}, draw opacity={0.25}, line width={0}, solid, mark={*}, mark size={3.75 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.25}, fill={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, fill opacity={0.25}, line width={0.75}, rotate={0}, solid}, only marks, forget plot]
        coordinates {
            (105.0,0.05584966441816506)
            (102.0,0.04975238296092152)
            (109.0,0.05940020202500067)
            (89.0,2.294967031333799)
            (109.0,0.01674682323492209)
            (103.0,0.10720479779621425)
            (111.0,0.047378610574098125)
            (91.0,0.10379493048125163)
            (110.0,2.513624607833005)
            (104.0,0.05086478924822935)
            (109.0,0.11351618280207543)
            (108.0,0.4020253123624037)
            (110.0,0.07446102987541804)
            (105.0,0.11535032606585913)
            (111.0,0.18480906070431535)
            (98.0,0.20276032629175944)
            (107.0,0.08951349573643062)
            (102.0,0.04128707134536484)
            (111.0,0.08564751808207169)
            (101.0,0.3024196229705965)
            (134.0,0.0625994120602839)
            (123.0,0.07725887987692295)
            (137.0,0.0446213228734711)
            (133.0,0.3084188318399655)
            (133.0,0.01962266981062221)
            (134.0,0.10146607016990984)
            (138.0,0.06814314041036643)
            (126.0,0.1400125295960749)
            (138.0,467.4741755918939)
            (127.0,1.1166477896146412)
            (138.0,0.17714972309111646)
            (19.0,0.11922132709056152)
            (133.0,0.08769252229223219)
            (130.0,0.07793316024608679)
            (137.0,0.1718723416265335)
            (125.0,0.33914255094342527)
            (136.0,0.015348260495236497)
            (131.0,0.14565843998621403)
            (137.0,0.19446231157170504)
            (123.0,0.26589194567164826)
            (178.0,0.13602372744489552)
            (164.0,1.3096147548790622)
            (183.0,1.2001412178607553)
            (20.0,0.13777438440294185)
            (181.0,0.8479907542370899)
            (183.0,0.0950066129025234)
            (182.0,0.0973834969835291)
            (140.0,4.6045371868841265)
            (181.0,0.08489426529555413)
            (171.0,1.434207428800683)
            (182.0,0.03715339077893962)
            (172.0,0.22976469267521113)
            (179.0,0.021032926937894693)
            (174.0,0.12032695402812911)
            (183.0,0.1184679992516246)
            (167.0,0.19684607920197908)
            (179.0,0.07686867922650574)
            (173.0,2.5546282892181007)
            (182.0,2.568218715361108)
            (20.0,0.14201835288630793)
            (57.0,0.034130835764939665)
            (57.0,0.08573571106663361)
            (57.0,0.027468464787432298)
            (53.0,0.1845821255465333)
            (57.0,0.3145059229030894)
            (50.0,0.03924953982860379)
            (57.0,0.06856875980583756)
            (18.0,0.13408208779207065)
            (57.0,0.246120844440766)
            (57.0,0.04960174895763792)
            (57.0,2.4305662054492216e9)
            (49.0,0.19845525108383108)
            (57.0,0.10124427200601582)
            (57.0,723.2519398047787)
            (57.0,0.04868286562065639)
            (21.0,0.09182349525167144)
            (57.0,2.72184197085509)
            (57.0,0.02883792885224287)
            (57.0,0.05463115607163071)
            (47.0,3.4065463877724946)
            (83.0,0.0897939767464811)
            (73.0,0.13931869624662485)
            (82.0,0.2953775450777107)
            (78.0,0.18845400004232993)
            (83.0,0.03326786171772812)
            (81.0,0.03346088666911155)
            (84.0,0.021920164977564584)
            (29.0,0.13107584979633108)
            (83.0,0.025448977666601213)
            (83.0,0.052182247335789286)
            (84.0,0.03318537664315938)
            (19.0,0.11324273930491076)
            (83.0,0.049563710942017516)
            (82.0,0.05229073956836107)
            (84.0,0.936285586965232)
            (79.0,0.36352072822414827)
            (84.0,0.06743210142176875)
            (83.0,0.06069929189699635)
            (84.0,1.0402239808527012)
            (20.0,0.0863532735177328)
        }
        ;
    \addplot[color={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, name path={0b6b36d8-c846-4873-8ad0-9609ab2cf7aa}, draw opacity={0.25}, line width={0}, solid, mark={pentagon*}, mark size={3.75 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.25}, fill={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, fill opacity={0.25}, line width={0.75}, rotate={0}, solid}, only marks, forget plot]
        coordinates {
            (235.0,0.13357198945321846)
            (70.0,0.04461548817109118)
            (240.0,0.27826514621649645)
            (55.0,0.40610599113961626)
            (231.0,0.16842724844486784)
            (183.0,0.053126061241543406)
            (243.0,0.09279724868377123)
            (44.0,0.6219680679576621)
            (224.0,0.14673187421056033)
            (222.0,0.17126303810452492)
            (242.0,0.26732925740823366)
            (72.0,0.40406046468282036)
            (239.0,0.27881283308599)
            (215.0,0.08349271759735011)
            (242.0,0.2125101399226415)
            (32.0,0.5513286633867127)
            (235.0,0.22568337408398828)
            (206.0,0.09258343306049661)
            (243.0,0.20454040997137593)
            (26.0,0.5107297023521244)
            (323.0,0.3667902032609065)
            (106.0,0.04378510564953422)
            (346.0,0.16065746518529136)
            (41.0,0.4591112638861731)
            (338.0,0.16645620302038866)
            (329.0,0.0837772625628679)
            (343.0,0.14366152756083547)
            (43.0,0.6024946038819587)
            (331.0,0.18929036413010172)
            (287.0,0.33185148527249725)
            (348.0,0.04576725492168219)
            (40.0,0.5307735589972514)
            (333.0,0.25561466836673213)
            (267.0,0.08139063563463969)
            (348.0,0.2603533867117459)
            (42.0,0.485145599462851)
            (324.0,0.06458173214154204)
            (150.0,0.03822594080269291)
            (346.0,0.22772148021684693)
            (38.0,0.5349911168200644)
            (529.0,0.09246980203670312)
            (76.0,0.032150726839285076)
            (561.0,0.2953731709925806)
            (42.0,0.42435150858456305)
            (551.0,0.5658710399734437)
            (172.0,0.03671331385133575)
            (562.0,0.2247420272307935)
            (39.0,0.6093866677344313)
            (552.0,0.42285681275210707)
            (338.0,0.07585812329891815)
            (561.0,0.16916001531239544)
            (39.0,0.4801687797714871)
            (536.0,0.1460284576306947)
            (519.0,0.1674664372593216)
            (563.0,0.07034107539784748)
            (36.0,0.6153149633129895)
            (538.0,0.2901270824759899)
            (314.0,0.0894523538965804)
            (561.0,0.9493069230558913)
            (43.0,0.45231290229037063)
            (86.0,0.19904079876071654)
            (86.0,0.2474597369029254)
            (87.0,0.9661924012140722)
            (52.0,0.5183448216596367)
            (83.0,0.41163416709756745)
            (86.0,0.2018533637471545)
            (87.0,0.7338740990737738)
            (29.0,0.502255804939499)
            (85.0,0.24871677122634128)
            (79.0,0.47124624050445574)
            (87.0,0.3017981172481904)
            (34.0,0.4118665138102874)
            (86.0,0.46110233661130057)
            (81.0,0.34380752518935775)
            (87.0,0.592914605387025)
            (55.0,0.5340532094284938)
            (84.0,0.21982864921620568)
            (86.0,0.12593689360335356)
            (86.0,0.2490899738548067)
            (30.0,0.3888134391760499)
            (149.0,0.155251101357749)
            (131.0,0.055439107403504)
            (155.0,0.12238978551321686)
            (35.0,0.43098325658284026)
            (154.0,0.1394374093998332)
            (152.0,0.2297799184953769)
            (155.0,0.15759694151794004)
            (39.0,0.5057258031699502)
            (154.0,0.3922494485884269)
            (153.0,0.3902229654498584)
            (156.0,0.2773388179874096)
            (43.0,0.5416488514754519)
            (152.0,0.2540652840489718)
            (152.0,0.11832827918668122)
            (154.0,0.11410277669310671)
            (45.0,0.5469697488849031)
            (151.0,0.1870216675213049)
            (127.0,0.16707232055495033)
            (154.0,0.8019823908524971)
            (52.0,0.5113011726127454)
        }
        ;
    \addplot[color={rgb,1:red,1.0;green,0.498;blue,0.0549}, name path={32500e26-f355-48fb-b27a-d77c0e6f6f11}, draw opacity={0.25}, line width={0}, solid, mark={diamond*}, mark size={3.75 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.25}, fill={rgb,1:red,1.0;green,0.498;blue,0.0549}, fill opacity={0.25}, line width={0.75}, rotate={0}, solid}, only marks, forget plot]
        coordinates {
            (75.0,3.147945165083429)
            (73.0,0.7787622330875555)
            (75.0,14.274696485077706)
            (72.0,1.0547043767865791)
            (75.0,3.332024981183407)
            (73.0,0.8054592789691524)
            (75.0,1.9717984455866802)
            (30.0,0.3567713595345774)
            (74.0,1.3257338505548062)
            (71.0,0.454412627271946)
            (75.0,199.32322467686308)
            (75.0,1.0895748631691455)
            (75.0,2.4040986461550835)
            (75.0,30.945783537221924)
            (75.0,1.3655781266539022)
            (72.0,1.7771060355248247)
            (75.0,2.095095365942199)
            (74.0,1.2957150765737162)
            (75.0,10.287194966343153)
            (71.0,23.823930630622364)
            (91.0,2.4310322835821725)
            (90.0,2.03837213281791)
            (93.0,18.48678688820552)
            (18.0,0.1828119186366187)
            (93.0,1.402802143535598)
            (93.0,0.5216804362331351)
            (93.0,344.95675624956914)
            (17.0,0.028392317708999436)
            (93.0,0.5279417818890894)
            (92.0,2.279971972737263)
            (93.0,1.6171895486111396)
            (87.0,0.9929110131500897)
            (92.0,1.6970363177778334)
            (88.0,1.281239828889679)
            (93.0,1.9682814650085763)
            (90.0,1.0443588384128824)
            (93.0,1.4672964427212034)
            (93.0,2.512126815083123)
            (93.0,1.7528157502014407e27)
            (17.0,0.042776379040499156)
            (123.0,292782.19801080116)
            (114.0,0.8319858220508034)
            (122.0,38.36679799718143)
            (119.0,4.247901992170559)
            (123.0,2.1603952895878047)
            (123.0,1.612179389285352)
            (123.0,2.1315469624808055e8)
            (120.0,2.863381885496416)
            (123.0,1.5701579511267307)
            (115.0,1.1995883679712909)
            (122.0,0.8442486879600574)
            (107.0,1.4996647965557675)
            (123.0,43.27286005493181)
            (123.0,0.22240221552099876)
            (123.0,1.724556189521123)
            (123.0,0.7600881434640848)
            (123.0,1.55953539967718)
            (122.0,2.5016970428713385)
            (122.0,125.83729920965449)
            (119.0,1.474402129759871)
            (39.0,1.6787168071085867)
            (38.0,23.286011001150722)
            (39.0,1.8524442120647987e6)
            (39.0,1.1819811023887494)
            (38.0,25.653816244212006)
            (39.0,7.724620004876016)
            (39.0,1.5998738392072849)
            (39.0,230.55843153298235)
            (38.0,35.384317652871445)
            (39.0,18.021070189844696)
            (39.0,195.47077720096976)
            (39.0,160.86609736256958)
            (39.0,3.0592725712776643)
            (39.0,4.398153803012362)
            (39.0,1.1590218432565333)
            (17.0,0.035424442595674274)
            (38.0,4162.440925088575)
            (39.0,40.27926776330599)
            (39.0,2.6373813625877043)
            (39.0,9.740619004230942)
            (57.0,0.8097482769396408)
            (57.0,0.8683241511330079)
            (57.0,0.6548577323997611)
            (57.0,18.271264624639596)
            (57.0,0.8110250229754907)
            (57.0,1.0118191812875524e19)
            (57.0,1166.0668083720273)
            (54.0,24.255595170101138)
            (57.0,2.4197291464377434e17)
            (57.0,25.939711745446342)
            (57.0,411.43443310327984)
            (54.0,25.852674031372572)
            (57.0,1.0362054731778492)
            (57.0,1.0072691225003283e14)
            (56.0,62.4551625376836)
            (47.0,0.3358689620078998)
            (57.0,2.1657544482083853)
            (57.0,1.253110778534678)
            (57.0,2.6027300344473447)
            (57.0,26.29584591974179)
        }
        ;
    \addplot[color={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, name path={17fa3149-1072-43c8-8c21-8a953ad9b789}, draw opacity={1.0}, line width={2}, dashed]
        coordinates {
            (26.0,0.5107297023521244)
            (30.0,0.3888134391760499)
            (70.0,0.04461548817109118)
            (76.0,0.032150726839285076)
            (172.0,0.03671331385133575)
            (348.0,0.04576725492168219)
            (563.0,0.07034107539784748)
        }
        ;
    \addlegendentry {Dense}
    \addplot[color={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, name path={92814b74-5dc8-4049-b104-86ec2005b545}, draw opacity={1.0}, line width={2}, solid]
        coordinates {
            (57.0,2.4305662054492216e9)
            (18.0,0.13408208779207065)
            (20.0,0.0863532735177328)
            (57.0,0.027468464787432298)
            (109.0,0.01674682323492209)
            (136.0,0.015348260495236497)
            (179.0,0.021032926937894693)
        }
        ;
    \addlegendentry {NPU}
    \addplot[color={rgb,1:red,1.0;green,0.498;blue,0.0549}, name path={01e229a2-0f53-4a30-b602-5c368ec01212}, draw opacity={1.0}, line width={2}, dashdotted]
        coordinates {
            (39.0,1.8524442120647987e6)
            (17.0,0.042776379040499156)
            (17.0,0.028392317708999436)
            (123.0,0.22240221552099876)
        }
        ;
    \addlegendentry {NPU (real)}
\end{axis}
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={SIR Model}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(0.98, 0.5)}, anchor={east}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={115.3mm}, yshift={-1.0mm}, width={112.3mm}, height={74.2mm}, scaled x ticks={false}, xlabel={Time}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmajorgrids={true}, xmin={-6.0}, xmax={206.0}, xtick={{0.0,50.0,100.0,150.0,200.0}}, xticklabels={{$0$,$50$,$100$,$150$,$200$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymajorgrids={true}, ymin={-3.0}, ymax={103.0}, ytick={{0.0,25.0,50.0,75.0,100.0}}, yticklabels={{$0$,$25$,$50$,$75$,$100$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={2c7c1a19-783e-45b8-ba0c-5f0d47ee61d7}, draw opacity={0.8}, line width={4}, solid]
        coordinates {
            (0.0,100.0)
            (5.128205128205129,97.46721309713212)
            (10.256410256410257,91.07411529978752)
            (15.384615384615385,81.85822003827998)
            (20.512820512820515,70.93406989207153)
            (25.641025641025642,59.36171599653322)
            (30.76923076923077,48.05121932732702)
            (35.8974358974359,37.70386572153941)
            (41.02564102564103,28.786661655619767)
            (46.15384615384615,21.536458098202502)
            (51.282051282051285,15.984367444258964)
            (56.41025641025641,12.000464352759552)
            (61.53846153846154,9.344211354396093)
            (66.66666666666667,7.730064928992786)
            (71.7948717948718,6.871778028659456)
            (76.92307692307692,6.531712440790295)
            (82.05128205128206,6.527846168986343)
            (87.17948717948718,6.732261015727868)
            (92.3076923076923,7.058906461697017)
            (97.43589743589743,7.448852247665159)
            (102.56410256410257,7.8621381580828835)
            (107.6923076923077,8.271957039163652)
            (112.82051282051282,8.660383314615434)
            (117.94871794871794,9.015599814052901)
            (123.07692307692308,9.330728378580337)
            (128.2051282051282,9.603063253817359)
            (133.33333333333334,9.832194944069075)
            (138.46153846153845,10.019480104740795)
            (143.5897435897436,10.168236969594094)
            (148.71794871794873,10.282830242207533)
            (153.84615384615384,10.367552891075434)
            (158.97435897435898,10.42665962038892)
            (164.10256410256412,10.464984481530161)
            (169.23076923076923,10.487122611260277)
            (174.35897435897436,10.496809113024415)
            (179.48717948717947,10.49696822418185)
            (184.6153846153846,10.490486949159731)
            (189.74358974358975,10.479876278697317)
            (194.87179487179486,10.467010848152167)
            (200.0,10.453187053405765)
        }
        ;
    \addlegendentry {True $S,I,R$}
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={1bd494f5-4860-4f91-bdd7-b6e623424164}, draw opacity={0.8}, line width={4}, solid, forget plot]
        coordinates {
            (0.0,0.01)
            (5.128205128205129,2.326002254590468)
            (10.256410256410257,7.535412858135904)
            (15.384615384615385,14.122344350839555)
            (20.512820512820515,20.861212946319004)
            (25.641025641025642,26.852061199698944)
            (30.76923076923077,31.52079010396936)
            (35.8974358974359,34.589940517583635)
            (41.02564102564103,36.03247808932137)
            (46.15384615384615,36.0106783095002)
            (51.282051282051285,34.81461042142043)
            (56.41025641025641,32.799670718853136)
            (61.53846153846154,30.321886606977262)
            (66.66666666666667,27.70444740533231)
            (71.7948717948718,25.186917212993873)
            (76.92307692307692,22.920167986169474)
            (82.05128205128206,20.97478194912876)
            (87.17948717948718,19.361879697078066)
            (92.3076923076923,18.06136663645284)
            (97.43589743589743,17.0351828644295)
            (102.56410256410257,16.243460467487697)
            (107.6923076923077,15.64716970581006)
            (112.82051282051282,15.209576700999039)
            (117.94871794871794,14.900102701702954)
            (123.07692307692308,14.692768418941395)
            (128.2051282051282,14.564262720391902)
            (133.33333333333334,14.495696710475123)
            (138.46153846153845,14.472242253795276)
            (143.5897435897436,14.480798749993546)
            (148.71794871794873,14.510674957112856)
            (153.84615384615384,14.553801666023665)
            (158.97435897435898,14.60424306180992)
            (164.10256410256412,14.656998699052162)
            (169.23076923076923,14.708526692892267)
            (174.35897435897436,14.75644891382746)
            (179.48717948717947,14.799483648325703)
            (184.6153846153846,14.836795847017735)
            (189.74358974358975,14.868211842226707)
            (194.87179487179486,14.893874068806547)
            (200.0,14.91412562301017)
        }
        ;
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={99fd42b1-24fe-476a-9976-bd874af9786a}, draw opacity={0.8}, line width={4}, solid, forget plot]
        coordinates {
            (0.0,0.0)
            (5.128205128205129,0.21678464827744337)
            (10.256410256410257,1.4004718420766002)
            (15.384615384615385,4.029435610880507)
            (20.512820512820515,8.214717161609485)
            (25.641025641025642,13.796222803767884)
            (30.76923076923077,20.43799056870365)
            (35.8974358974359,27.71619376087699)
            (41.02564102564103,35.19086025505889)
            (46.15384615384615,42.462863592297325)
            (51.282051282051285,49.21102213432063)
            (56.41025641025641,55.20986492838734)
            (61.53846153846154,60.34390203862667)
            (66.66666666666667,64.57548766567494)
            (71.7948717948718,67.95130475834671)
            (76.92307692307692,70.55811957304027)
            (82.05128205128206,72.50737188188494)
            (87.17948717948718,73.9158592871941)
            (92.3076923076923,74.88972690185018)
            (97.43589743589743,75.52596488790537)
            (102.56410256410257,75.90440137442945)
            (107.6923076923077,76.09087325502632)
            (112.82051282051282,76.14003998438557)
            (117.94871794871794,76.09429748424418)
            (123.07692307692308,75.9865032024783)
            (128.2051282051282,75.84267402579079)
            (133.33333333333334,75.68210834545584)
            (138.46153846153845,75.51827764146397)
            (143.5897435897436,75.36096428041239)
            (148.71794871794873,75.21649480067964)
            (153.84615384615384,75.08864544290094)
            (158.97435897435898,74.97909731780119)
            (164.10256410256412,74.8880168194177)
            (169.23076923076923,74.81435069584748)
            (174.35897435897436,74.75674197314815)
            (179.48717948717947,74.71354812749247)
            (184.6153846153846,74.68271720382256)
            (189.74358974358975,74.661911879076)
            (194.87179487179486,74.64911508304131)
            (200.0,74.6426873235841)
        }
        ;
    \addplot[color={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, name path={78950b23-4587-48e7-ac69-7a173b42d7ff}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,100.0)
            (5.128205128205129,97.06371891032103)
            (10.256410256410257,91.16684462650328)
            (15.384615384615385,82.31860705364566)
            (20.512820512820515,71.37328028612147)
            (25.641025641025642,59.56655329185587)
            (30.76923076923077,48.02982853680665)
            (35.8974358974359,37.56283706154292)
            (41.02564102564103,28.637211640920665)
            (46.15384615384615,21.450945359433096)
            (51.282051282051285,15.991646351608207)
            (56.41025641025641,12.091156352995037)
            (61.53846153846154,9.484490451837605)
            (66.66666666666667,7.875556914992195)
            (71.7948717948718,6.990870233379205)
            (76.92307692307692,6.607983321954687)
            (82.05128205128206,6.5600222783518465)
            (87.17948717948718,6.726996153400294)
            (92.3076923076923,7.024919064760524)
            (97.43589743589743,7.39470727063654)
            (102.56410256410257,7.794775985610217)
            (107.6923076923077,8.196178604628553)
            (112.82051282051282,8.579336543097728)
            (117.94871794871794,8.931747895191842)
            (123.07692307692308,9.246307992466145)
            (128.2051282051282,9.520025067810243)
            (133.33333333333334,9.752717909881044)
            (138.46153846153845,9.946211973122255)
            (143.5897435897436,10.103649973188396)
            (148.71794871794873,10.228936301333658)
            (153.84615384615384,10.326310541375907)
            (158.97435897435898,10.400033334390523)
            (164.10256410256412,10.454186099682712)
            (169.23076923076923,10.492511634963721)
            (174.35897435897436,10.518286052520692)
            (179.48717948717947,10.534341485628904)
            (184.6153846153846,10.543067590956252)
            (189.74358974358975,10.546434143372755)
            (194.87179487179486,10.546027918810093)
            (200.0,10.54309740549285)
        }
        ;
    \addlegendentry {$\hat S$}
    \addplot[color={rgb,1:red,1.0;green,0.498;blue,0.0549}, name path={55448a83-940d-42d2-bfa5-a69e99f44840}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,1.0;green,0.498;blue,0.0549}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,0.01)
            (5.128205128205129,2.7125599630590216)
            (10.256410256410257,7.66090322551522)
            (15.384615384615385,14.05992089552267)
            (20.512820512820515,20.853377037596527)
            (25.641025641025642,26.949263303743713)
            (30.76923076923077,31.643268521967897)
            (35.8974358974359,34.65445060429321)
            (41.02564102564103,36.00248493995587)
            (46.15384615384615,35.89734765592843)
            (51.282051282051285,34.65263524249474)
            (56.41025641025641,32.62658545846643)
            (61.53846153846154,30.171255479960887)
            (66.66666666666667,27.588258250205744)
            (71.7948717948718,25.101967545018532)
            (76.92307692307692,22.854404251300224)
            (82.05128205128206,20.916574933981376)
            (87.17948717948718,19.306505476039106)
            (92.3076923076923,18.00963924091414)
            (97.43589743589743,16.993337294186258)
            (102.56410256410257,16.217140779474875)
            (107.6923076923077,15.639502083735772)
            (112.82051282051282,15.2216338588409)
            (117.94871794871794,14.929362607150622)
            (123.07692307692308,14.733677548787432)
            (128.2051282051282,14.610599644370984)
            (133.33333333333334,14.541013265661315)
            (138.46153846153845,14.509771044572922)
            (143.5897435897436,14.505057091871647)
            (148.71794871794873,14.517816285691135)
            (153.84615384615384,14.541245312235125)
            (158.97435897435898,14.570351813286242)
            (164.10256410256412,14.601580037074164)
            (169.23076923076923,14.632505354378589)
            (174.35897435897436,14.661561456708021)
            (179.48717948717947,14.687819457461144)
            (184.6153846153846,14.710818905225636)
            (189.74358974358975,14.730432472745509)
            (194.87179487179486,14.746759324518035)
            (200.0,14.760042741576434)
        }
        ;
    \addlegendentry {$\hat I$}
    \addplot[color={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, name path={7029124d-f96e-4c96-8ccc-df7e7adb01de}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,0.0)
            (5.128205128205129,0.50693326871289)
            (10.256410256410257,1.417679355799354)
            (15.384615384615385,3.8182491876122566)
            (20.512820512820515,7.962302503151513)
            (25.641025641025642,13.649388710326537)
            (30.76923076923077,20.427370387367066)
            (35.8974358974359,27.787946129597962)
            (41.02564102564103,35.263551206919445)
            (46.15384615384615,42.46863602460634)
            (51.282051282051285,49.116287047058584)
            (56.41025641025641,55.02090402143257)
            (61.53846153846154,60.09164848079044)
            (66.66666666666667,64.31440131855449)
            (71.7948717948718,67.72832791710991)
            (76.92307692307692,70.40495950573064)
            (82.05128205128206,72.43348328533506)
            (87.17948717948718,73.91170893256104)
            (92.3076923076923,74.93627783745409)
            (97.43589743589743,75.5987646759572)
            (102.56410256410257,75.98223718673248)
            (107.6923076923077,76.15882412766837)
            (112.82051282051282,76.18868881450685)
            (117.94871794871794,76.12017219498782)
            (123.07692307692308,75.99077303726688)
            (128.2051282051282,75.8285172837301)
            (133.33333333333334,75.65344697694208)
            (138.46153846153845,75.47937800364895)
            (143.5897435897436,75.31528402015716)
            (148.71794871794873,75.16647568233809)
            (153.84615384615384,75.03557984442344)
            (158.97435897435898,74.92332819239051)
            (164.10256410256412,74.82915576306749)
            (169.23076923076923,74.75168343757129)
            (174.35897435897436,74.68913253538693)
            (179.48717948717947,74.63952887626489)
            (184.6153846153846,74.60087490919402)
            (189.74358974358975,74.57126520497009)
            (194.87179487179486,74.54895766923063)
            (200.0,74.53241178440281)
        }
        ;
    \addlegendentry {$\hat R$}
\end{axis}
\end{tikzpicture}