### 2020-12-17 slides_short/fsir.tex
more slides

\begin{tikzpicture}[/tikz/background rectangle/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, draw opacity={1.0}}, show background rectangle]
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, title={fSIR Model}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, at={(0.98, 0.5)}, anchor={east}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={115.3mm}, yshift={-1.0mm}, width={112.3mm}, height={74.2mm}, scaled x ticks={false}, xlabel={Time}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmajorgrids={true}, xmin={-6.0}, xmax={206.0}, xtick={{0.0,50.0,100.0,150.0,200.0}}, xticklabels={{$0$,$50$,$100$,$150$,$200$}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis x line*={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymajorgrids={true}, ymin={-3.0}, ymax={103.0}, ytick={{0.0,25.0,50.0,75.0,100.0}}, yticklabels={{$0$,$25$,$50$,$75$,$100$}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, axis y line*={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}]
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={4d2d32cf-0078-477b-aca5-ac47612e3b35}, draw opacity={0.8}, line width={4}, solid]
        coordinates {
            (0.0,100.0)
            (5.128205128205129,97.46721309713212)
            (10.256410256410257,91.07411529978752)
            (15.384615384615385,81.85822003827998)
            (20.512820512820515,70.93406989207153)
            (25.641025641025642,59.36171599653322)
            (30.76923076923077,48.05121932732702)
            (35.8974358974359,37.70386572153941)
            (41.02564102564103,28.786661655619767)
            (46.15384615384615,21.536458098202502)
            (51.282051282051285,15.984367444258964)
            (56.41025641025641,12.000464352759552)
            (61.53846153846154,9.344211354396093)
            (66.66666666666667,7.730064928992786)
            (71.7948717948718,6.871778028659456)
            (76.92307692307692,6.531712440790295)
            (82.05128205128206,6.527846168986343)
            (87.17948717948718,6.732261015727868)
            (92.3076923076923,7.058906461697017)
            (97.43589743589743,7.448852247665159)
            (102.56410256410257,7.8621381580828835)
            (107.6923076923077,8.271957039163652)
            (112.82051282051282,8.660383314615434)
            (117.94871794871794,9.015599814052901)
            (123.07692307692308,9.330728378580337)
            (128.2051282051282,9.603063253817359)
            (133.33333333333334,9.832194944069075)
            (138.46153846153845,10.019480104740795)
            (143.5897435897436,10.168236969594094)
            (148.71794871794873,10.282830242207533)
            (153.84615384615384,10.367552891075434)
            (158.97435897435898,10.42665962038892)
            (164.10256410256412,10.464984481530161)
            (169.23076923076923,10.487122611260277)
            (174.35897435897436,10.496809113024415)
            (179.48717948717947,10.49696822418185)
            (184.6153846153846,10.490486949159731)
            (189.74358974358975,10.479876278697317)
            (194.87179487179486,10.467010848152167)
            (200.0,10.453187053405765)
        }
        ;
    \addlegendentry {True $S,I,R$}
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={b32b3689-d920-4aa7-9895-f3013f9b1a1b}, draw opacity={0.8}, line width={4}, solid, forget plot]
        coordinates {
            (0.0,0.01)
            (5.128205128205129,2.326002254590468)
            (10.256410256410257,7.535412858135904)
            (15.384615384615385,14.122344350839555)
            (20.512820512820515,20.861212946319004)
            (25.641025641025642,26.852061199698944)
            (30.76923076923077,31.52079010396936)
            (35.8974358974359,34.589940517583635)
            (41.02564102564103,36.03247808932137)
            (46.15384615384615,36.0106783095002)
            (51.282051282051285,34.81461042142043)
            (56.41025641025641,32.799670718853136)
            (61.53846153846154,30.321886606977262)
            (66.66666666666667,27.70444740533231)
            (71.7948717948718,25.186917212993873)
            (76.92307692307692,22.920167986169474)
            (82.05128205128206,20.97478194912876)
            (87.17948717948718,19.361879697078066)
            (92.3076923076923,18.06136663645284)
            (97.43589743589743,17.0351828644295)
            (102.56410256410257,16.243460467487697)
            (107.6923076923077,15.64716970581006)
            (112.82051282051282,15.209576700999039)
            (117.94871794871794,14.900102701702954)
            (123.07692307692308,14.692768418941395)
            (128.2051282051282,14.564262720391902)
            (133.33333333333334,14.495696710475123)
            (138.46153846153845,14.472242253795276)
            (143.5897435897436,14.480798749993546)
            (148.71794871794873,14.510674957112856)
            (153.84615384615384,14.553801666023665)
            (158.97435897435898,14.60424306180992)
            (164.10256410256412,14.656998699052162)
            (169.23076923076923,14.708526692892267)
            (174.35897435897436,14.75644891382746)
            (179.48717948717947,14.799483648325703)
            (184.6153846153846,14.836795847017735)
            (189.74358974358975,14.868211842226707)
            (194.87179487179486,14.893874068806547)
            (200.0,14.91412562301017)
        }
        ;
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={d04156b5-29d1-4f76-8661-f226866e6ee2}, draw opacity={0.8}, line width={4}, solid, forget plot]
        coordinates {
            (0.0,0.0)
            (5.128205128205129,0.21678464827744337)
            (10.256410256410257,1.4004718420766002)
            (15.384615384615385,4.029435610880507)
            (20.512820512820515,8.214717161609485)
            (25.641025641025642,13.796222803767884)
            (30.76923076923077,20.43799056870365)
            (35.8974358974359,27.71619376087699)
            (41.02564102564103,35.19086025505889)
            (46.15384615384615,42.462863592297325)
            (51.282051282051285,49.21102213432063)
            (56.41025641025641,55.20986492838734)
            (61.53846153846154,60.34390203862667)
            (66.66666666666667,64.57548766567494)
            (71.7948717948718,67.95130475834671)
            (76.92307692307692,70.55811957304027)
            (82.05128205128206,72.50737188188494)
            (87.17948717948718,73.9158592871941)
            (92.3076923076923,74.88972690185018)
            (97.43589743589743,75.52596488790537)
            (102.56410256410257,75.90440137442945)
            (107.6923076923077,76.09087325502632)
            (112.82051282051282,76.14003998438557)
            (117.94871794871794,76.09429748424418)
            (123.07692307692308,75.9865032024783)
            (128.2051282051282,75.84267402579079)
            (133.33333333333334,75.68210834545584)
            (138.46153846153845,75.51827764146397)
            (143.5897435897436,75.36096428041239)
            (148.71794871794873,75.21649480067964)
            (153.84615384615384,75.08864544290094)
            (158.97435897435898,74.97909731780119)
            (164.10256410256412,74.8880168194177)
            (169.23076923076923,74.81435069584748)
            (174.35897435897436,74.75674197314815)
            (179.48717948717947,74.71354812749247)
            (184.6153846153846,74.68271720382256)
            (189.74358974358975,74.661911879076)
            (194.87179487179486,74.64911508304131)
            (200.0,74.6426873235841)
        }
        ;
    \addplot[color={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, name path={daf90f21-20cf-4381-9c0f-30de95c3d54b}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.1216;green,0.4667;blue,0.7059}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,100.0)
            (5.128205128205129,97.06371891032103)
            (10.256410256410257,91.16684462650328)
            (15.384615384615385,82.31860705364566)
            (20.512820512820515,71.37328028612147)
            (25.641025641025642,59.56655329185587)
            (30.76923076923077,48.02982853680665)
            (35.8974358974359,37.56283706154292)
            (41.02564102564103,28.637211640920665)
            (46.15384615384615,21.450945359433096)
            (51.282051282051285,15.991646351608207)
            (56.41025641025641,12.091156352995037)
            (61.53846153846154,9.484490451837605)
            (66.66666666666667,7.875556914992195)
            (71.7948717948718,6.990870233379205)
            (76.92307692307692,6.607983321954687)
            (82.05128205128206,6.5600222783518465)
            (87.17948717948718,6.726996153400294)
            (92.3076923076923,7.024919064760524)
            (97.43589743589743,7.39470727063654)
            (102.56410256410257,7.794775985610217)
            (107.6923076923077,8.196178604628553)
            (112.82051282051282,8.579336543097728)
            (117.94871794871794,8.931747895191842)
            (123.07692307692308,9.246307992466145)
            (128.2051282051282,9.520025067810243)
            (133.33333333333334,9.752717909881044)
            (138.46153846153845,9.946211973122255)
            (143.5897435897436,10.103649973188396)
            (148.71794871794873,10.228936301333658)
            (153.84615384615384,10.326310541375907)
            (158.97435897435898,10.400033334390523)
            (164.10256410256412,10.454186099682712)
            (169.23076923076923,10.492511634963721)
            (174.35897435897436,10.518286052520692)
            (179.48717948717947,10.534341485628904)
            (184.6153846153846,10.543067590956252)
            (189.74358974358975,10.546434143372755)
            (194.87179487179486,10.546027918810093)
            (200.0,10.54309740549285)
        }
        ;
    \addlegendentry {$\hat S$}
    \addplot[color={rgb,1:red,1.0;green,0.498;blue,0.0549}, name path={d719ac49-0b45-4546-885d-0971ecba6177}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,1.0;green,0.498;blue,0.0549}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,0.01)
            (5.128205128205129,2.7125599630590216)
            (10.256410256410257,7.66090322551522)
            (15.384615384615385,14.05992089552267)
            (20.512820512820515,20.853377037596527)
            (25.641025641025642,26.949263303743713)
            (30.76923076923077,31.643268521967897)
            (35.8974358974359,34.65445060429321)
            (41.02564102564103,36.00248493995587)
            (46.15384615384615,35.89734765592843)
            (51.282051282051285,34.65263524249474)
            (56.41025641025641,32.62658545846643)
            (61.53846153846154,30.171255479960887)
            (66.66666666666667,27.588258250205744)
            (71.7948717948718,25.101967545018532)
            (76.92307692307692,22.854404251300224)
            (82.05128205128206,20.916574933981376)
            (87.17948717948718,19.306505476039106)
            (92.3076923076923,18.00963924091414)
            (97.43589743589743,16.993337294186258)
            (102.56410256410257,16.217140779474875)
            (107.6923076923077,15.639502083735772)
            (112.82051282051282,15.2216338588409)
            (117.94871794871794,14.929362607150622)
            (123.07692307692308,14.733677548787432)
            (128.2051282051282,14.610599644370984)
            (133.33333333333334,14.541013265661315)
            (138.46153846153845,14.509771044572922)
            (143.5897435897436,14.505057091871647)
            (148.71794871794873,14.517816285691135)
            (153.84615384615384,14.541245312235125)
            (158.97435897435898,14.570351813286242)
            (164.10256410256412,14.601580037074164)
            (169.23076923076923,14.632505354378589)
            (174.35897435897436,14.661561456708021)
            (179.48717948717947,14.687819457461144)
            (184.6153846153846,14.710818905225636)
            (189.74358974358975,14.730432472745509)
            (194.87179487179486,14.746759324518035)
            (200.0,14.760042741576434)
        }
        ;
    \addlegendentry {$\hat I$}
    \addplot[color={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, name path={209e1cdb-c150-414a-bfc5-63b07ba91f51}, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.1725;green,0.6275;blue,0.1725}, fill opacity={1.0}, line width={0.75}, rotate={0}, solid}, only marks]
        coordinates {
            (0.0,0.0)
            (5.128205128205129,0.50693326871289)
            (10.256410256410257,1.417679355799354)
            (15.384615384615385,3.8182491876122566)
            (20.512820512820515,7.962302503151513)
            (25.641025641025642,13.649388710326537)
            (30.76923076923077,20.427370387367066)
            (35.8974358974359,27.787946129597962)
            (41.02564102564103,35.263551206919445)
            (46.15384615384615,42.46863602460634)
            (51.282051282051285,49.116287047058584)
            (56.41025641025641,55.02090402143257)
            (61.53846153846154,60.09164848079044)
            (66.66666666666667,64.31440131855449)
            (71.7948717948718,67.72832791710991)
            (76.92307692307692,70.40495950573064)
            (82.05128205128206,72.43348328533506)
            (87.17948717948718,73.91170893256104)
            (92.3076923076923,74.93627783745409)
            (97.43589743589743,75.5987646759572)
            (102.56410256410257,75.98223718673248)
            (107.6923076923077,76.15882412766837)
            (112.82051282051282,76.18868881450685)
            (117.94871794871794,76.12017219498782)
            (123.07692307692308,75.99077303726688)
            (128.2051282051282,75.8285172837301)
            (133.33333333333334,75.65344697694208)
            (138.46153846153845,75.47937800364895)
            (143.5897435897436,75.31528402015716)
            (148.71794871794873,75.16647568233809)
            (153.84615384615384,75.03557984442344)
            (158.97435897435898,74.92332819239051)
            (164.10256410256412,74.82915576306749)
            (169.23076923076923,74.75168343757129)
            (174.35897435897436,74.68913253538693)
            (179.48717948717947,74.63952887626489)
            (184.6153846153846,74.60087490919402)
            (189.74358974358975,74.57126520497009)
            (194.87179487179486,74.54895766923063)
            (200.0,74.53241178440281)
        }
        ;
    \addlegendentry {$\hat R$}
\end{axis}
\end{tikzpicture}